Apparent preferences for cognitive effort fade when multiple forms of effort

### 2021-03-04 Cost_manuscript.Rmd
new latex for models

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort
  and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey^1,2^, Gary A. Kane^1,2^, & Joseph T. McGuire^1,2^"
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
- \usepackage{indentfirst}
output:
  pdf_document:
    highlight: haddock
    keep_tex: no
  word_document: #default
    reference_docx: Style_reference_draft.docx
  html_document:
    df_print: paged
csl: apa.csl
indent: true
bibliography: cost_paper.bib
---

@@

## figures at the end of the manuscropt?
figsEnd <- FALSE

@@

$$\begin{aligned} 
\dfrac {\sum p_i R_i} {Time_{offer} + (\sum p_i H + Time_{reward}) + 3T},p \in \{0, 1\}
\end{aligned}$$

@@

\begin{align}
  Model_{base} = R_i - \gamma H_i
\end{align}

@@

\begin{align}
  P(accept)_i = \dfrac {1} {1 + e^{-(1/\beta)(Model_{base})}}
\end{align}

@@

\begin{align}
  Model_{adaptive} = R_i - \gamma_i H_i
\end{align}

@@

\begin{align}
  Model_{biased} = R_i - (\gamma + mC_i) H_i
\end{align}

@@

\begin{align}
  Model_{adaptive + biased} = R_i - (\gamma_i + mC_i) H_i
\end{align}

@@

``` {r BTW: 16.1.4. Optimality (move to acceptance comparison eventually), echo = FALSE}
# in order
# clean data, 
# calculate each individual's prop. accept,
# calculate t.tests with optimality as null mean,
# correct pvalues with FDR
temp <- dataBtw %>%
  filter(Choice < 2) %>%
  group_by(SubjID, Handling, Offer, Cost) %>%
  summarise(pAccept = mean(Choice)) %>%
  mutate(optimal = case_when(
          (Handling == 10 & Offer < 8) ~ 0,
          (Handling == 14 & Offer < 20) ~ 0,
          TRUE ~ 1
        )) %>% 
  group_by(Handling, Offer, Cost) %>%
  summarise(meanAccept = mean(pAccept),
            sdAccept = sd(pAccept),
            optimal = unique(optimal),
            pvals = tryCatch(t.test(pAccept, mu = unique(optimal))$p.value, error = function(e) {1})) %>%
  mutate(FDR = p.adjust(pvals, method = "BY"),
         significant = FDR < 0.05)

# proportion of significant ones
# eventually find a way to plot this (or just place it further down)
# alternative per block: temp %>% group_by(Handling) %>% summarise(prop = mean(significant, na.rm = T))
optimalProp <- mean(temp$significant, na.rm = T) * 100

# proportion accepted per cost condition
optimalProp_cost <- temp %>% 
  group_by(Cost) %>% 
  summarise(propDeviations = round(mean(significant) * 100))

# ## if you want to replicate the cost 3 method
# # this says that all but easy deviated from optimality in the observed direction
# temp <- dataBtw %>%
#   group_by(Cost) %>%
#   filter(Offer < 20, Handling == 10) %>%
#   summarize(propAccept = mean(Choice),
#             accept = sum(Choice),
#             quit = sum(Choice == 0))
# 
# temp2 <- apply(temp, 1, function(row) {prop.test(as.numeric(row[3]), sum(as.numeric(row[3:4])), p = 0.5, conf.level = 0.95)})
#   
#   
# # store stats in a dataframe
# news <- data.frame(probability = sapply(temp2, "[[", "estimate"),
#                                    chisquared = sapply(temp2, "[[", "statistic"),
#                                    confint = t(sapply(temp2, "[[", "conf.int")),
#                                    pval = sapply(temp2, "[[", "p.value"))
# rownames(news) <- temp$Cost

@@

  We next examined the optimality of these decisions. The foraging task was configured so that there was a single reward-maximizing strategy per block type (see Operationalization of Cost in Methods), and participants were warned that accepting everything would not maximize their gains. Accordingly, participants who over- or under-accepted earned the least (Figure 3, right; general linear model with quadratic term, F = `r round(summary(lmEarn)$fstatistic[1], digits = 2)`, Beta = `r round(summary(lmEarn)$coefficients[3,1], digits = 2)`, SE = `r round(summary(lmEarn)$coefficients[3, 2], digits = 2)`, $R^2$ = `r round(summary(lmEarn)$r.squared, digits = 2)`, *p* `r report_p(summary(lmEarn)$coefficients[3, 4])`). Two-sided, one-sample t-tests against the optimal proportion of acceptances (i.e. $\mu$ = 0 or 1) for each combination of cost type, handling time, and offer showed that participants significantly deviated from optimality in `r round(optimalProp)`% of combinations (*p* < 0.05, FDR corrected), with the percentage of deviations being lowest for the wait group (Wait = `r filter(optimalProp_cost, Cost == "Wait")[2]`%, physical = `r filter(optimalProp_cost, Cost == "Physical")[2]`%, Cognitive = `r filter(optimalProp_cost, Cost == "Cognitive")[2]`%, Easy = `r filter(optimalProp_cost, Cost == "Easy")[2]`%). Together, these findings show that even though participants did not always perform optimally, they adequately integrated the statistics of the environment as they tracked their recent experience. Further, they suggest that cognitive effort can boost the value of an offer, as reflected by higher acceptances at the expense of earnings.

@@

include_graphics("./Images/FIG1_short.png")
```

@@

include_graphics("./Images/FIG2.png")
```

>>>>

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort
  and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey^1,2^, Gary A. Kane^1,2^, & Joseph T. McGuire^1,2^"
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
- \usepackage{indentfirst}
- \usepackage{fixltx2e}
output:
  pdf_document:
    highlight: haddock
    keep_tex: no
  word_document: #default
    reference_docx: Style_reference_draft.docx
  html_document:
    df_print: paged
csl: apa.csl
indent: true
bibliography: cost_paper.bib
---

@@

## figures at the end of the manuscript?
figsEnd <- TRUE

@@

$$\begin{aligned} 
g(*) = \dfrac {\sum p_i R_i} { 3\text{Time}_\text{offer} + (\sum p_i H + \text{Time}_\text{reward}) + 3T},p \in \{0, 1\}
\end{aligned}$$

@@

\begin{align}
  \text{Model}_\text{base} = R_i - \gamma H_i
\end{align}

@@

\begin{align}
  P(\text{accept})_i = \dfrac {1} {1 + e^{-(1/\beta)(\text{Model}_\text{base})}}
\end{align}

@@

\begin{align}
  \text{Model}_\text{adaptive} = R_i - \gamma_i H_i
\end{align}

@@

\begin{align}
  \text{Model}_\text{biased} = R_i - (\gamma + mC_i) H_i
\end{align}

@@

\begin{align}
  \text{Model}_\text{adaptive + biased} = R_i - (\gamma_i + mC_i) H_i
\end{align}

@@

``` {r BTW: 16.1.4. Optimality (move to acceptance comparison eventually), echo = FALSE}
# # in order
# # clean data, 
# # calculate each individual's prop. accept,
# # calculate t.tests with optimality as null mean,
# # correct pvalues with FDR
# temp <- dataBtw %>%
#   filter(Choice < 2) %>%
#   group_by(SubjID, Handling, Offer, Cost) %>%
#   summarise(pAccept = mean(Choice)) %>%
#   mutate(optimal = case_when(
#           (Handling == 10 & Offer < 8) ~ 0,
#           (Handling == 14 & Offer < 20) ~ 0,
#           TRUE ~ 1
#         )) %>% 
#   group_by(Handling, Offer, Cost) %>%
#   summarise(meanAccept = mean(pAccept),
#             sdAccept = sd(pAccept),
#             optimal = unique(optimal),
#             pvals = tryCatch(t.test(pAccept, mu = unique(optimal))$p.value, error = function(e) {1})) %>%
#   mutate(FDR = p.adjust(pvals, method = "BY"),
#          significant = FDR < 0.05)
# 
# 
# # proportion of significant ones
# # eventually find a way to plot this (or just place it further down)
# # alternative per block: temp %>% group_by(Handling) %>% summarise(prop = mean(significant, na.rm = T))
# optimalProp <- mean(temp$significant, na.rm = T) * 100
# 
# # proportion accepted per cost condition
# optimalProp_cost <- temp %>% 
#   group_by(Cost) %>% 
#   summarise(propDeviations = round(mean(significant) * 100))

## if you want to replicate the cost 3 method
# this says that all but easy deviated from optimality in the observed direction
temp <- dataBtw %>%
  group_by(Cost) %>%
  filter(Offer < 20) %>%
  summarize(propAccept = mean(Choice),
            accept = sum(Choice),
            quit = sum(Choice == 0))

temp2 <- apply(temp, 1, function(row) {prop.test(as.numeric(row[3]), sum(as.numeric(row[3:4])), p = 0.5, conf.level = 0.95)})

# store stats in a dataframe
optimalProp_cost <- data.frame(probability = sapply(temp2, "[[", "estimate"),
                                   chisquared = sapply(temp2, "[[", "statistic"),
                                   confint = t(sapply(temp2, "[[", "conf.int")),
                                   pval = sapply(temp2, "[[", "p.value"))
rownames(optimalProp_cost) <- temp$Cost

@@

  We next examined the optimality of these decisions. The foraging task was configured so that there was a single reward-maximizing strategy per block type (see Operationalization of Cost in Methods), and participants were warned that accepting everything would not maximize their gains. Accordingly, participants who over- or under-accepted earned the least (Figure 3, right; general linear model with quadratic term, F = `r round(summary(lmEarn)$fstatistic[1], digits = 2)`, Beta = `r round(summary(lmEarn)$coefficients[3,1], digits = 2)`, SE = `r round(summary(lmEarn)$coefficients[3, 2], digits = 2)`, $R^2$ = `r round(summary(lmEarn)$r.squared, digits = 2)`, *p* `r report_p(summary(lmEarn)$coefficients[3, 4])`). We performed group-level chi-squared tests of proportions to examine whether each cost biased decision makers to under- or over-accept 4 and 8 cent offers, which together should have been accepted 50% of the time. We found a significant over-acceptance of these trials by the cognitive effort group (proportion accepted = `r round(optimalProp_cost["Cognitive", "probability"], digits = 2)`%; $\chi^2$ = `r round(optimalProp_cost["Cognitive", "chisquared"], digits = 2)`, *p* `r report_p(summary_costvsreward["Cognitive", "pval"]`), but under-acceptances by the physical (proportion accepted = `r round(optimalProp_cost["Physical", "probability"], digits = 2)`%; $\chi^2$ = `r round(optimalProp_cost["Physical", "chisquared"], digits = 2)`, *p* `r report_p(summary_costvsreward["Physical", "pval"]`) and wait (proportion accepted = `r round(optimalProp_cost["Physical", "probability"], digits = 2)`%; $\chi^2$ = `r round(optimalProp_cost["Physical", "chisquared"], digits = 2)`, *p* `r report_p(summary_costvsreward["Physical", "pval"]`) groups (no significant difference for the Easy group). Together, these findings show that even though participants did not always perform optimally, they adequately integrated the statistics of the environment as they tracked their recent experience. Further, they suggest that cognitive effort can boost the value of an offer, as reflected by higher acceptances at the expense of earnings.

@@

@@
